---
output: reprex::reprex_document
knit: reprex::reprex_render
---

# Data Analysis: Self-Paced Reading Task

## Research Questions and Predictions

**Research questions**:

1. Are object relative clauses (ORs) parsed slower than subject relative clauses (SRs) by adult Spanish-English bilinguals?

2. If so, does number mismatch between the two determiner phrases (DPs) facilitate their comprehension?

**Predictions**:

1. Adult Spanish-English bilinguals will show shorter reaction times (RTs) in SRs than in ORs.

2. Number mismatch between the two DPs will result in shorter RTs in ORs than number match.

## Data Analysis

This is a step by step guide on how the data collected by the self-paced reading task has been analyzed.

### 1. Library Loading

The following libraries are imported:

1. `tidyverse`, for cleaner data wrangling functions and access to other included packages, such as `ggplot2`.
2. `ggthemes`, to access the theme `theme_calc()`.
3. `lmerTest`, for more powerful mixed regression models.
4. `gt`, for table creation functions.

```{r}
library(tidyverse)
library(ggthemes)
library(lmerTest)
library(gt)
```

### 2. Importing and Processing Data
First, an example is taken to take a look at the state of the data and the existing variables. 

```{r}
df_example <- read.csv("./data/32.csv")
summary(df_example)
```