---
title: "Self-Paced Reading Task"
output: html_document
--- 

## Data Analysis

### 0. Setting Up
#### Loading Libraries
```{r Importing Libraries, results=FALSE, message=FALSE}
library(tidyverse)
library(lmerTest)
library(ggthemes)
library(gt)
library(broom.mixed)
```

#### Color Variables
```{r Setting Color Variables}
accent_color <- "#354c7c"
secondary_color <- "#D5A419"
```

#### Path Variables
```{r Setting Path Variables}
tables_path <- "./images/tables/"
plots_path <- "./images/plots/"
```

#### Setting DefaultPlot Theme
```{r Setting Default Plot Theme}
theme_set(
  theme_calc(
    base_size = 12,
    base_family = "Times"
  ) +
    theme(
      plot.title = element_text(size = 12, face = "bold", margin = margin(b = 10)),
      plot.subtitle = element_text(size = 11, face = "bold"),
      plot.background = element_rect(linewidth = 0),
      strip.background = element_rect(fill = accent_color, color = accent_color),
      strip.text = element_text(color = "white", face = "bold"),
      axis.title.x = element_text(face = "bold", margin = margin(t = 10)),
      axis.title.y = element_text(face = "bold", margin = margin(r = 10)),
      axis.line = element_line(color = accent_color, linewidth = 1),
      axis.ticks = element_blank(),
      legend.title = element_text(face = "bold")
    )
)
```

### 1. Participant Data
#### Importing Participant Data
```{r Importing Participant Data}
df_part <- read.csv("./data/part_data.csv")
df_part$part_name <- NULL
df_part$part_mail <- NULL
head(df_part)
```

#### Cleaning Participant Data
```{r Cleaning Participant Data}
df_part <- df_part %>%
  select(part_experiment, part_id, part_age, part_lang, part_impaired., part_abroad, part_certificate, part_cat_exp, part_cat_pro, part_spa_exp, part_spa_pro, part_eng_exp, part_eng_pro, part_eng_onset_time) %>%
  mutate(
    across(
      c(part_experiment, part_id, part_age, part_lang, part_impaired., part_abroad, part_certificate, part_cat_exp, part_cat_pro, part_spa_exp, part_spa_pro, part_eng_exp, part_eng_pro),
      factor
    )
  ) %>%
  filter(part_experiment == "TFG") %>%
  rename(part_id = part_id) %>%
  droplevels()

df_part$part_experiment <- NULL

summary(df_part)
```

### 2. Task Data
#### Importing Task Data
```{r Importing Example of Task Data}
df_example <- read.csv("./data/32.csv")
head(df_example)
```

```{r Importing Task Data}
df_raw <- reduce(
  list(
    select(read.csv("./data/29.csv"), part_id, task, item_number, item_type, clause_type, number, segment, rt),
    select(read.csv("./data/31.csv"), part_id, task, item_number, item_type, clause_type, number, segment, rt),
    select(read.csv("./data/32.csv"), part_id, task, item_number, item_type, clause_type, number, segment, rt),
    select(read.csv("./data/37.csv"), part_id, task, item_number, item_type, clause_type, number, segment, rt)
  ),
  rbind
)
glimpse(df_raw)
```

#### Cleaning Task Data
```{r Cleaning Task Data}
df_clean <- df_raw %>%
  arrange(part_id) %>%
  filter(.$item_type == "exp_item" & .$task == "item" & segment != "s0") %>%
  mutate(
    across(
      c(task, item_type, part_id),
      factor
    ),
    item_number = as.numeric(item_number),
    clause_type = factor(clause_type, levels = c("sr", "or"), labels = c("SR", "OR")),
    number = factor(number, levels = c("ss", "pp", "sp", "ps"), labels = c("SS", "PP", "SP", "PS")),
    matching = factor(ifelse(number == "SS" | number == "PP", "match", "mismatch"), labels = c("Match", "Mismatch")),
    segment = factor(segment, labels = c("S1", "S2", "S3", "S4", "S5")),
  ) %>%
  relocate(matching, .after = number) %>%
  droplevels()

glimpse(df_clean)
```

#### Normalizing Task Data
```{r Normality Check}
skewed_plot <- df_clean %>%
  ggplot(aes(x = rt)) +
  geom_histogram(aes(y = ..density..)) +
  stat_function(
    fun = dnorm,
    args = list(mean = mean(df_clean$rt), sd = sd(df_clean$rt)),
    linewidth = 1,
    color = accent_color
  )
skewed_plot
```

```{r Logged Data}
df_clean <- df_clean %>%
  mutate(rt = log(rt))

normal_plot <- df_clean %>%
  ggplot(aes(x = rt)) +
  geom_histogram(aes(y = ..density..)) +
  stat_function(
    fun = dnorm,
    args = list(mean = mean(df_clean$rt), sd = sd(df_clean$rt)),
    linewidth = 1,
    color = accent_color
  )
normal_plot
```

```{r Outlier Removal}
q1 <- quantile(df_clean$rt, 0.25)
q3 <- quantile(df_clean$rt, 0.75)
iqr <- q3 - q1
lower_limit <- q1 - iqr * 1.5
upper_limit <- q3 + iqr * 1.5

exp(lower_limit)
exp(upper_limit)

n_obs_pre <- nrow(df_clean)

df_clean <- df_clean %>%
  filter(.$rt > lower_limit & .$rt < upper_limit)

n_obs_post <- nrow(df_clean)

n_outliers <- n_obs_pre - n_obs_post
n_outliers

n_outliers * 100 / n_obs_pre
```

```{r Distribution Plot}
distribution_plot <- df_clean %>%
  ggplot(aes(x = rt)) +
  geom_histogram(
    aes(y = ..density..),
    fill = secondary_color
  ) +
  stat_function(
    fun = dnorm,
    args = list(mean = mean(df_clean$rt), sd = sd(df_clean$rt)),
    linewidth = 1,
    color = accent_color
  ) +
  labs(
    title = "RT Data Distribution",
    x = "Reaction Times (log10)",
    y = "Density"
  )
distribution_plot

ggsave(filename = "distribution_plot.svg", path = plots_path, height = 10, width = 15, unit = "cm")
```

```{r Normality Test}
summary(df_clean$rt)
sd(df_clean$rt)

shapiro.test(df_clean$rt)

qq_plot <- df_clean %>%
  ggplot(aes(sample = rt)) +
  stat_qq() +
  stat_qq_line(color = accent_color, linewidth = 1) +
  labs(title = "Q-Q Plot of RTs", x = "Theoretical Quantiles", y = "Sample Quantiles")
qq_plot
```

#### Merging Participant and Task Data
```{r Participant and Test Data Merging}
df_part$part_experiment <- NULL

df_merged <- merge(df_clean, df_part, by = "part_id")
summary(df_merged)
```

### 3. Mixed Effects Models
#### Clause Effects Model
```{r Clause Effects Model}
lm_clause <- df_merged %>%
  filter(segment %in% c("S3", "S4", "S5")) %>%
  lmer(data = ., rt ~ clause_type + (1 | part_eng_pro) + (1 | part_eng_exp) + (1 | part_id))
summary(lm_clause)
```

#### Clause Effects Table
```{r Clause Effects Clean Output}
lm_clause_tidy <- lm_clause %>%
  tidy() %>%
  filter(effect == "fixed") %>%
  mutate(
    estimate = exp(estimate),
    code = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*",
      p.value > 0.05 ~ " "
    ),
    p.value = case_when(
      p.value < 0.001 ~ "< 0.001",
      p.value < 0.01 ~ "< 0.01",
      p.value < 0.05 ~ "< 0.05",
      p.value > 0.05 ~ as.character(round(p.value, 2))
    ),
    term = case_when(
      term == "(Intercept)" ~ "SR (Baseline)",
      term == "clause_typeOR" ~ "OR"
    )
  )

lm_clause_tidy
```

```{r Clause Effects Table}
lm_clause_table <- lm_clause_tidy %>%
  gt() %>%
  cols_hide(columns = c(effect, group, df)) %>%
  tab_options(
    table.width = 565,
    table.font.names = "Times",
    table.font.size = 12,
    table.border.top.width = 0,
    heading.title.font.size = 12,
    heading.subtitle.font.size = 11,
    heading.border.bottom.color = accent_color,
    column_labels.font.weight = "bold",
    column_labels.background.color = accent_color
  ) %>%
  tab_header(
    title = md("**Mixed Model Results**"),
    subtitle = md("**Clause Type**")
  ) %>%
  cols_label(
    term = "Condition",
    estimate = "Estimate (ms)",
    std.error = "Std. Error",
    statistic = md("*t*-value"),
    p.value = md("*p*-value"),
    code = "Signif. Code"
  ) %>%
  fmt_number(
    columns = c(term, estimate, std.error, statistic, p.value),
    decimals = 3
  )

lm_clause_table
gtsave(lm_clause_table, filename = "clause_type_table.png", path = tables_path)
```

#### Clause Effects Plot
```{r Clause Effects Plot}
lm_clause_plot <- df_merged %>%
  filter(segment %in% c("S3", "S4")) %>%
  ggplot(
    aes(
      y = exp(rt),
      x = clause_type,
      color = clause_type,
      fill = clause_type
    )
  ) +
  geom_boxplot(
    show.legend = FALSE,
    staplewidth = 0.5,
    notch = TRUE,
    alpha = 0.5,
    linewidth = 0.5
  ) +
  geom_rug(
    show.legend = FALSE,
    sides = "l"
  ) +
  labs(
    title = "Clause Type Effects on RC RTs",
    subtitle = "Data from Segments 3 and 4",
    x = "Clause Type",
    y = "Reaction Times (ms)"
  )
lm_clause_plot

ggsave(filename = "lm_clause_plot.svg", path = plots_path, height = 10, width = 10, unit = "cm")
```

#### Number Effects Model
```{r Number Effects Model}
lm_number <- df_merged %>%
  filter(segment %in% c("S3", "S4", "S5")) %>%
  lmer(data = ., rt ~ clause_type * number + (1 | part_eng_exp) + (1 | part_eng_pro))
summary(lm_number)
```

#### Number Effects Table
```{r Number Effects Clean Output}
lm_number_tidy <- lm_number %>%
  tidy() %>%
  filter(effect == "fixed") %>%
  mutate(
    estimate = exp(estimate),
    code = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*",
      p.value > 0.05 ~ " "
    ),
    p.value = case_when(
      p.value < 0.001 ~ "< 0.001",
      p.value < 0.01 ~ "< 0.01",
      p.value < 0.05 ~ "< 0.05",
      p.value > 0.05 ~ as.character(round(p.value, 2)),
    ),
    term = case_when(
      term == "(Intercept)" ~ "SR * SS (Baseline)",
      term == "clause_typeOR" ~ "OR",
      term == "numberPP" ~ "PP",
      term == "numberSP" ~ "SP",
      term == "numberPS" ~ "PS",
      term == "clause_typeOR:numberPP" ~ "OR * PP",
      term == "clause_typeOR:numberSP" ~ "OR * SP",
      term == "clause_typeOR:numberPS" ~ "OR * PS",
    )
  )

lm_number_tidy
```

```{r Number Effects Table}
lm_number_table <- lm_number_tidy %>%
  gt() %>%
  cols_hide(columns = c(effect, group, df)) %>%
  tab_options(
    table.width = 565,
    table.font.names = "Times",
    table.font.size = 12,
    table.border.top.width = 0,
    heading.title.font.size = 12,
    heading.subtitle.font.size = 11,
    column_labels.font.weight = "bold",
    heading.border.bottom.color = accent_color,
    column_labels.background.color = accent_color
  ) %>%
  tab_header(
    title = md("**Mixed Model Results**"),
    subtitle = md("**Number and ORs**")
  ) %>%
  cols_label(
    term = "Condition",
    estimate = "Estimate (ms)",
    std.error = "Std. Error",
    statistic = md("*t*-value"),
    p.value = md("*p*-value"),
    code = "Signif. Code"
  ) %>%
  fmt_number(
    columns = c(term, estimate, std.error, statistic, p.value),
    decimals = 3
  )

lm_number_table
gtsave(lm_number_table, filename = "number_table.png", path = tables_path)
```

#### Number Effects Plot
```{r Number Effects Plot}
lm_number_plot <- df_merged %>%
  filter(clause_type == "OR") %>%
  ggplot(
    aes(
      y = exp(rt),
      x = segment,
      color = number,
      group = number
    )
  ) +
  stat_summary(
    fun = mean,
    geom = "line",
    linewidth = 1
  ) +
  stat_summary(
    aes(shape = number),
    fun = mean,
    geom = "point",
    size = 2,
    stroke = 1,
    fill = "white"
  ) +
  geom_rug(
    show.legend = FALSE,
    sides = "l",
    alpha = 0.5
  ) +
  coord_cartesian(ylim = c(600, 2100)) +
  scale_shape_manual(
    values = c(21, 22, 23, 24)
  ) +
  scale_x_discrete(
    labels = c(
      "S1\nThe dog/s",
      "S2\nthat",
      "S3\nthe cat/s",
      "S4\nchase/s",
      "S5\nrun/s"
    )
  ) +
  labs(
    title = "Number Effect on OR RTs by Segment",
    x = "Segments",
    y = "Reaction Times (ms)"
  ) +
  guides(
    color = guide_legend("Number"),
    shape = guide_legend("Number")
  ) +
  theme(
    legend.position = c(0.11, 0.29),
    legend.box.background = element_rect(linewidth = 1, color = accent_color, linetype = "solid"),
    legend.text = element_text(size = 11),
    legend.title = element_text(size = 11)
  ) +
  facet_wrap(~matching, dir = "v")
lm_number_plot

ggsave(filename = "lm_number_segment_plot.svg", path = plots_path, height = 15, width = 15, units = "cm")
```
